Measurement Bias from Address Aliasing

### 2014-11-20 paper.tex
Added PDF meta data
Replace paragraph with subsubsection
Paragraph did not render very well, and subsubsection is the correct
next level in the hierarchy

% paper.tex
% ACM sigmetric format
% Author: L. K. Melhus and R. E. Jensen
\documentclass{acm_proc_article-sp}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{tikz}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{underscore}
\usepackage{url}
\usepackage{lmodern} % Much clearer text
\usepackage[utf8]{inputenc}

@@

@@

\paragraph{Note on Hyper-Threading}
As described in Section~\ref{sec:methodology}, Hyper-Threading was disabled by default in order to avoid any potential interference while collecting data.
However, we found that configuration of Hyper-Threading either way did not change the alias behavior.
As an example of this, we show the results for ``Nehalem'' with HT enabled in Figure~\ref{fig:microkernel-all-archs}.

@@

\paragraph{Mark buffers with \texttt{restrict}}
In our convolution kernel implementation, the compiler has to account for the fact that input and output pointers might alias, or that the buffers partially overlap.
This limits the extent generated code can keep data in registers without updating the values from memory, as a write to one buffer potentially could invalidate a cached value from the other.
The C99 keyword \emph{restrict} can be used to explicitly tell the compiler that accesses through a pointer does not alias with any other, allowing for more efficient code generation with fewer memory accesses.
\begin{lstlisting}[breaklines=true]
    void conv(int n, const float * restrict input, float * restrict output);
\end{lstlisting}
With this updated function prototype, the compiler is able to move a store instruction out of the inner loop, reducing the number of stores executed by more than 80 \% on optimization level O2.
Alias events is reduced by about 10 million for the default alignment, with a corresponding improvement in cycle count. % File backing this in results

\paragraph{Use a special purpose allocator}
The Intel optimization manual mentions this in \emph{User/Source Coding Rule 8}, suggesting that special purpose allocators could be used to avoid aliasing ~\cite{OptimizationManual}.
However, to our knowledge there are no commonly used allocators that specifically try to mitigate aliasing in heap allocated memory.
%Similarly none offset their allocations based on the initial stack frame.
We show that heap allocators are prone to generate pairwise aliasing buffers, the worst case for functions such as the convolution example.

@@

\paragraph{Manually adjust address offsets}
In some cases it might make sense to explicitly control the memory addresses used, for example forcing some fixed relative offset between input and output pointers.
This can be achieved by exploiting the fact the \texttt{mmap} is in fact guaranteed to be placed at a page boundary, and use that directly instead of \texttt{malloc}.
The following approach can be used to make an anonymous memory mapping with offset \texttt{d} bytes away from page alignment.
\begin{lstlisting}[breaklines=true]
    mmap(NULL, (n + d), PROT_READ | PROT_WRITE, MAP_PRIVATE | MAP_ANONYMOUS, -1, 0) + d;
\end{lstlisting}
The same pointer difference must of course by subtracted before unmapping the memory.

>>>>

% paper.tex
% ACM sigmetric format
% Author: L. K. Melhus and R. E. Jensen
\documentclass{acm_proc_article-sp}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{tikz}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{underscore}
\usepackage{url}
\usepackage{lmodern} % Much clearer text
\usepackage[utf8]{inputenc}
\usepackage{hyperref} % For PDF metadata

@@

\hypersetup
{
%    pdfauthor={L. K. Melhus and R. E. Jensen},
    pdfauthor={Anonymous and Anonymous},
    pdfsubject={Identification of an underlying mechanism that causes measurement bias},
    pdftitle={Measurement Bias from Address Aliasing},
    pdfkeywords={4K address aliasing, BLAS, heap allocators, measurement bias, memory disambiguation, performance counters}
}

@@

\subsubsection{Note on Hyper-Threading}
As described in Section~\ref{sec:methodology}, Hyper-Threading was disabled by default in order to avoid any potential interference while collecting data.
However, we found that configuration of Hyper-Threading either way did not change the alias behavior.
As an example of this, we show the results for ``Nehalem'' with HT enabled in Figure~\ref{fig:microkernel-all-archs}.

@@

\subsubsection{Mark buffers with \texttt{restrict}}
In our convolution kernel implementation, the compiler has to account for the fact that input and output pointers might alias, or that the buffers partially overlap.
This limits the extent generated code can keep data in registers without updating the values from memory, as a write to one buffer potentially could invalidate a cached value from the other.
The C99 keyword \emph{restrict} can be used to explicitly tell the compiler that accesses through a pointer does not alias with any other, allowing for more efficient code generation with fewer memory accesses.
\begin{lstlisting}[breaklines=true]
    void conv(int n, const float * restrict input, float * restrict output);
\end{lstlisting}
With this updated function prototype, the compiler is able to move a store instruction out of the inner loop, reducing the number of stores executed by more than 80 \% on optimization level O2.
Alias events is reduced by about 10 million for the default alignment, with a corresponding improvement in cycle count. % File backing this in results

\subsubsection{Use a special purpose allocator}
The Intel optimization manual mentions this in \emph{User/Source Coding Rule 8}, suggesting that special purpose allocators could be used to avoid aliasing ~\cite{OptimizationManual}.
However, to our knowledge there are no commonly used allocators that specifically try to mitigate aliasing in heap allocated memory.
%Similarly none offset their allocations based on the initial stack frame.
We show that heap allocators are prone to generate pairwise aliasing buffers, the worst case for functions such as the convolution example.

@@

\subsubsection{Manually adjust address offsets}
In some cases it might make sense to explicitly control the memory addresses used, for example forcing some fixed relative offset between input and output pointers.
This can be achieved by exploiting the fact the \texttt{mmap} is in fact guaranteed to be placed at a page boundary, and use that directly instead of \texttt{malloc}.
The following approach can be used to make an anonymous memory mapping with offset \texttt{d} bytes away from page alignment.
\begin{lstlisting}[breaklines=true]
    mmap(NULL, (n + d), PROT_READ | PROT_WRITE, MAP_PRIVATE | MAP_ANONYMOUS, -1, 0) + d;
\end{lstlisting}
The same pointer difference must of course by subtracted before unmapping the memory.
